Florida 2016 US Presidential Campaign Contribution Analysis
========================================================
by Xilin Miao

#Introduction
This is an exploration of 2016 US presidential campaigns financial contributions of Florida. Presidential campaign finance data contains much informatino wich is very valuable to reveal the demographic and geographic support for each party and candidate, and it also reflect the current status of the candidates in the race.

The reason why I chose Florida's data is that Florida has been the most pupulated of the "swing states", means where both political parties have a similar level of support. For this exploratory data analysis, the data of 2016 presidential campaigns financial contributions is from [Federal Election Commission](http://classic.fec.gov/disclosurep/pnational.do).

In this analysis, I will try to answer four questions: 
Firstly, which party and candidates received the most financial support in Florida (from the donations received)?
Secondly, is there a difference in donations between genders, and what is the patterns here?
Next, who made these contributions? which occupation make the most?
Lastly, where do these donations come from geographically?

#Dataset
Download dataset and load data.

```{r}
# Load the Data
FL <- read.csv("C:/Files/Udacity/Data-Analysis/Projects/Florida 2016 US Presidential Campaign Contribution Analysis/P00000001-FL.csv", stringsAsFactors = F, row.names = NULL)[,-19]
colnames(FL) <- c('cmte_id', 'cand_id', 'cand_nm', 'contbr_nm',	'contbr_city', 
                  'contbr_st', 'contbr_zip',	'contbr_employer', 
                  'contbr_occupation',	'contb_receipt_amt', 'contb_receipt_dt',
                  'receipt_desc', 'memo_cd', 'memo_text', 'form_tp', 'file_num', 
                  'tran_id', 'election_tp')

str(FL)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(ggthemes)
library(gridExtra)
library(dplyr)
library(tidyr)
library(stringr)
library(zipcode)
library(reshape2)
library(gender)
library(lubridate)
library(maps)
library(choroplethr)
library(choroplethrMaps)
library(choroplethrZip)
```


##Data Preprocessing
###Create New Variables
Before analyzing this data set, we should firstly do data processing since some features which I am interested in are not included in such as gender, and donor's geographical location in latigude and longtitude. 

```{r}
# Create party variable

democrat <- c("Clinton, Hillary Rodham",
                "Lessig, Lawrence",
                "O'Malley, Martin Joseph",
                "Sanders, Bernard",
                "Webb, James Henry Jr.")

others <- c("Johnson, Gary", "McMullin, Evan", "Stein, Jill")

FL$party <- ifelse(FL$cand_nm %in% democrat, "Democrat", "Republican")
FL$party[FL$cand_nm %in% others] <- 'Others'
```

```{r}
# Extract first name from contbr_nm for gender predication
FL$contbr_first_nm <- sub(" .*", "", sub(".*, ", "", FL$contbr_nm))
```

```{r}
#Create gender dataframe
gender_df <- gender(FL$contbr_first_nm, method = "ssa", c(1932, 2012), countries = "United States")  
# "ssa" here means looks up names based from the U.S Social Security Administration baby name data

#Create gender variable
gender_df <- unique(gender_df)
names(gender_df)[1] = "contbr_first_nm"
FL <- inner_join(FL, gender_df, by = 'contbr_first_nm')
```


```{r}
# Convert zipcode to latitude and longitude by zipcode package
data(zipcode)
FL$contbr_zip <- substr(FL$contbr_zip, 1, 5)
zipcode_FL <- subset(zipcode, state == "FL")[, -c(2,3)]
colnames(zipcode_FL) <- c("contbr_zip", "latitude", "longitude")
FL <- merge(FL, zipcode_FL)
```

```{r}
#Extract conbution recipt year and month
FL <- FL %>%
  mutate(date = as.Date(contb_receipt_dt, "%d-%b-%y"),
         year = year(date),
         month = month(date),
         year_month = paste(month.abb[month], ",", year))
```

###Data Clean
After looking into this data set, we noticed that there are 7608 negative contributions (contb_receipt_amt). These negative values should be the refund, so we should omit these observations. And since there is [contribution limits for 2015-2016](https://www.fec.gov/help-candidates-and-committees/candidate-taking-receipts/contribution-limits-candidates/), there is a limit which is $2,700 per election, per candidate. So the contribution above $2,700 will also be omitted since these contribution will be refunded lastly.

```{r}
nrow(FL[FL$contb_receipt_amt <= 0,])
```
```{r}
FL = filter(FL, FL$contb_receipt_amt > 0 & FL$contb_receipt_amt <= 2700)
```

##Dataset Overview
```{r}
str(FL)
```

After processing the data, we have removed observations with either negative amount or amount exceed $2,700. Except [the original 18 variables](ftp://ftp.fec.gov/FEC/Presidential_Map/2016/DATA_DICTIONARIES/CONTRIBUTOR_FORMAT.txt), following main variables were added also:
* party: candidate' party affiliation
* latitude: contributors' geographic latitude
* longitude: contributors' geographic lonitude
* contbr_first_nm: contributors' first name
* gender: contributors' gender
* date: contribution receipt data
* year: contribution receipt year
* month: contribution receipt month
* year_month: contribution receipt date (year month format)

#Univariate Plots Section
```{r}
summary(FL$contb_receipt_amt)
```

To start, let's have a glance of how the contribution distributed.
```{r}
p1 <- ggplot(aes(x = contb_receipt_amt), data = FL) +
  geom_histogram(bins = 50)
p2 <- ggplot(aes(x = 1, y = contb_receipt_amt), data = FL) +
  geom_boxplot()

grid.arrange(p1, p2, ncol = 2)
```

From upper plots we can see that there are so many outliers. To better understand the distribution of the contribution, now transform the plots to log10.
```{r}
ggplot(aes(x = contb_receipt_amt), data = FL) +
  geom_histogram(binwidth = 0.05) +
  scale_x_log10() +
  ggtitle('Histogram of the Contribution (with log10)')

tail(sort(table(FL$contb_receipt_amt)), 5)
summary(FL$contb_receipt_amt)
```

The distribution looks normal now and we can see that most donors made small amount of contributions which is among about 50 to 125. Now let's see the contribution distribution across the parties, candidates, genders and occupations.


```{r}
#Create party dataframe with Descending order
party_group <- group_by(FL, party)
FL.contr_by_party <- summarize(party_group,
                               sum_party = sum(contb_receipt_amt),
                               number_of_candidate = length(unique(cand_id)),
                               mean_party = sum_party/number_of_candidate,
                               n = n())

FL.contr_by_party[rev(order(FL.contr_by_party$n)),]

```

```{r}
#plot of party
ggplot(aes(x = reorder(party, -n), y = n, fill = party), data = FL.contr_by_party) +
  geom_bar(stat = 'identity') +
  geom_text(stat = 'identity', aes(label = n),
            data = FL.contr_by_party, vjust = -0.4) +
  xlab('Party') +
  ylab('Total Number of Contributions') +
  ggtitle('Total Number of Contributions by Party') +
  scale_fill_manual(values = c('blue', 'gold', 'red'))

sum(FL.contr_by_party$n)
```
From this we can see that the total number of 2016 presidential campaign finance contributions is near 405K, while the Democratic party took around 257K and almost 1.75 times of the number of donations make to the Republican party, which is 147K, and other parties took very few numbers of donations compared with Democrat party and Republican party, with about 1.3K. 


```{r}
#Create Candidate dataframe with Descending order
cand_group <- group_by(FL, cand_nm)
FL.contr_by_cand <- summarize(cand_group,
                               sum_cand = sum(contb_receipt_amt),
                               n = n())

FL.contr_by_cand[rev(order(FL.contr_by_cand$n)),]
```

```{r}
ggplot(aes(x = reorder(cand_nm, -n), y = n), data = FL.contr_by_cand) +
  geom_bar(stat = 'identity') +
  xlab('candidate') +
  ylab('Number of Contributions') +
  ggtitle('Number of Contributions by Candidate') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

There were intotal 25 candidates, and Clinton, Hillary Rodham took the most number of contributions with about 176.8K, and followed by Sanders, Bernard with about 79.6K and Trump, Donald J. with about 74.2K.

```{r}
#Create gender dataframe by descending order
gender_group <- group_by(FL, gender)
FL.contr_by_gender <- summarize(gender_group,
                               sum_gen = sum(contb_receipt_amt),
                               mean_gen = mean(contb_receipt_amt),
                               n = n())
FL.contr_by_gender[rev(order(FL.contr_by_gender$n)),]
```

```{r}
#Plot of gender
ggplot(aes(x = gender, y = n, fill = gender), 
       data = FL.contr_by_gender, vjust = -0.4) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = n), stat = 'identity', data = FL.contr_by_gender, vjust = -0.4) +
  xlab('Gender') +
  ylab('Number of Contributions') +
  ggtitle('Number of Contributions by Gender')
```

We can see here women and men made comparable donations, with about 199K from women and almost 206K from men. But from the mean we can see there is obvious difference. Let's see what are these donors occupation.

```{r}
#Create occupation dataframe with Descending order
occupation_group <- group_by(FL, contbr_occupation)
FL.contr_by_occupation <- summarize(occupation_group,
                               sum_occupation = sum(contb_receipt_amt),
                               mean_occupation = mean(contb_receipt_amt),
                               n = n())

#"INFORMATION REQUESTED" means the donor haven't provide this information, ignored
FL.contr_by_occupation <- subset(FL.contr_by_occupation, contbr_occupation != "INFORMATION REQUESTED" & contbr_occupation != "INFORMATION REQUESTED PER BEST EFFORTS")
FL.contr_by_occupation <- head(arrange(FL.contr_by_occupation,desc(n)), n = 10)
FL.contr_by_occupation[rev(order(FL.contr_by_occupation$n)),]
```

```{r}
#plot of occupation
ggplot(aes(x = reorder(contbr_occupation, -n), y = n), data = FL.contr_by_occupation) +
  geom_bar(stat = 'identity') +
  xlab('Top 10 Occupations') +
  ylab('Number of Donors') +
  ggtitle('Top 10 Occupations by Number of Donors') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

After looking into occupations, we can find that retired people take the first place, followed by not employed people, attorney comes to the third.

```{r}
summary(FL$contb_receipt_dt)
```

```{r}
#Create city dataframe by descending order
city_group <- group_by(FL, contbr_city)
FL.contr_by_city <- summarize(city_group,
                               sum_city = sum(contb_receipt_amt),
                               mean_city = mean(contb_receipt_amt),
                               n = n())
FL.contr_by_city <- head(arrange(FL.contr_by_city,desc(n)), n = 10)
FL.contr_by_city[rev(order(FL.contr_by_city$n)),]
```

```{r}
#plot of city
ggplot(aes(x = reorder(contbr_city, -n), y = n), data = FL.contr_by_city) +
  geom_bar(stat = 'identity') +
  xlab('Top 10 City') +
  ylab('Number of Donors') +
  ggtitle('Top 10 City by Number of Donors') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The city with most number of contribution is Miami, with about 23.6K.

```{r}
#plot of date
ggplot(aes(x = date), data = FL) +
  geom_histogram(position = position_dodge(), binwidth = 7) +
  xlab('Date') +
  ylab('Number of Donors') +
  ggtitle('Countribution Counts by Date')
```

It's interesting to see when people made contributions. this distribution shows that the bimodal with peaking around June 2016 and again when closing to the election.

# Univariate Analysis

### What is the structure of your dataset?
There are 404968 contributions and 18 variables. The variables that I am more interested in are:
* cand_nm: Candidate Name
* contbr_nm: Contributor name(first name, for gender prediction)
* contvr_zip:Contributor Zipcode
* contbr_occupation: COntributor Occupation
* contb_receipt_amt: COntribution Amount
* contb_receipt_dt: Contribution date

Other observations:
* The median contribution amount is $35.
* Most people contribute small amount of money, around 50 to 125.
* The numbers of contribution of women and men are comparable, while in average men's contribution amout is almost 1.5 times of women's.
* The Democrat received the most number of donations, and it is almost 1.75 times of Republican.
* Hillary Clinton have the most supporters.
* The city with most number of contribution is Miami.
* Retired people make the most number of contributions.

### What is/are the main feature(s) of interest in your dataset?
The main features of interest in this dataset are party, candidate, and contribution amount. I would like to find the answers to my questions at the beginning and also try to use combination of variables to predictive a donor's contribution party.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
Features like gender, contributor's occupation, contribution date, contributor's zipcode can help support further insights of the dataset. We can explore more by finding the relationships between these features and the main features of interest above.

### Did you create any new variables from existing variables in the dataset?
Yes, 5 new variables are created:
* party: candidates party.
* contbr_first_nm: contributor's first name to predict contributos' gender.
* gender: contributor's gender.
* latitude: contributors' geographic latitude.
* longitude: contributors' geographic lonitude.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
In this dataset, the observations with contribution amount value below 0 and 2700 were omitted since negative values are refund and amount value higher than 2700 will be refunded.

# Bivariate Plots Section
Let's see the barplots for contribution amount by party and candidates.

```{r}
FL.contr_by_party[rev(order(FL.contr_by_party$sum_party)),]
```

```{r}
#plot of contribution amount by party
ggplot(aes(x = reorder(party, -(sum_party/1000)), y = sum_party/1000, fill = party), data = FL.contr_by_party) +
  geom_bar(stat = 'identity') +
  geom_text(stat = 'identity', aes(label = round(sum_party/1000)),
            data = FL.contr_by_party, vjust = -0.4) +
  xlab('Party') +
  ylab('Contribution Amounts (Thousands)') +
  ggtitle('Total Contribution Amounts by Party') +
  scale_fill_manual(values = c('blue', 'gold', 'red'))
```

It's interesting to see shere the contribution money went. Previouly in the contribution number plot, we see Democrat took number almost 1.75 time of Republican. But now from the total amount, we can see Republican is leading now, with amount about $32.6K, almost 1.3 times of Democrat.

Since the nuber of candidates is quite different between Democrat and Republican, with 5 Democrat candidates while 17 for Republican candidates, let's see the average amount plot.

```{r}
#plot of contribution amount by party
ggplot(aes(x = reorder(party, -(mean_party/1000)), y = mean_party/1000, fill = party), data = FL.contr_by_party) +
  geom_bar(stat = 'identity') +
  geom_text(stat = 'identity', aes(label = round(mean_party/1000)),
            data = FL.contr_by_party, vjust = -0.4) +
  xlab('Party') +
  ylab('Contribution Amounts (Thousands)') +
  ggtitle('Average Contribution Amounts by Party') +
  scale_fill_manual(values = c('blue', 'gold', 'red'))
```

After looking at the average amount, we can see that now Democrat is almost 2.6 times of Republican in the average contribution amount.

Now let's see the contribution amount distribution by candidates.

```{r}
FL.contr_by_cand[rev(order(FL.contr_by_cand$sum_cand)),]
```

```{r}
#plot of contribution amount by candidates
ggplot(aes(x = reorder(cand_nm, -(sum_cand/1000)), y = sum_cand/1000), data = FL.contr_by_cand) +
  geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab('Candidate') +
  ylab('Contribution Amount (Thousands)') +
  ggtitle('Contribution Amount by Candidate')

sum(FL$contb_receipt_amt)
```

Same with the nuber of contributions, Hillary Cliton received the most contribution amount with about $21,451.7K, but followed by Donald Trump with $12,386K, Rubio Marco received about $6,634K as the third most contribution amount.

Now let's see the contribution patterns between parties and candidates.
```{r}
#boxplot of contribution amount by party
ggplot(aes(x = reorder(party, -contb_receipt_amt), y = contb_receipt_amt, fill = party), data = FL) +
  geom_boxplot() +
  xlab('party') +
  ylab('Contribution Amount') +
  ggtitle('Boxplot for Contribution Amount by Party') +
  scale_fill_manual(values = c('blue', 'gold', 'red'))
```

It's hard to compare contributions among all parties since there are too many outliers. And since now I am focuting on Demotrat party and Republican party (both contribution numbers and amount of other parties are not comparable with these 2 main party), the "Ohters" will be removed for now.

```{r}
#remove "Others"
FL <- subset(FL, FL$cand_nm != "Johnson, Gary" & FL$cand_nm != "McMullin, Evan" & FL$cand_nm != "Stein, Jill")

by(FL$contb_receipt_amt, FL$party, summary)
```


```{r}
#Add log10: boxplot of contribution amount by party
ggplot(aes(x = reorder(party, -contb_receipt_amt), y = contb_receipt_amt, fill = party), data = FL) +
  geom_boxplot() +
  scale_y_log10() +
  xlab('party') +
  ylab('Log10 Contribution Amount') +
  ggtitle('Boxplot for Contribution Amount by Party') +
  scale_fill_manual(values = c('blue', 'red'))
```

From this boxplot we can see that Republican party has higher median and mean, and Democrat party has more variations, that is contributions to Democrat party has more great and small amount of donations.

```{r}
#boxplot of contribution amount by candidates in descending order
by(FL$contb_receipt_amt, FL$cand_nm, summary)

ggplot(aes(x = reorder(cand_nm, -contb_receipt_amt), y = contb_receipt_amt), data = FL) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab('candidate') +
  ylab('Contribution Amount') +
  ggtitle('Contribution Amount by Candidate')
```

Remember earilier we saw Hillary Cliton received the most contribution amount and followed by Donald Trump and then Rubio Marco received the third most contribution amount. Now from the boxplot, Perry James R. (Rick) has the higherst median althought very few number and amount of contributions. Donald Trump and Hilary Clition seem to have very low median, but they both has a lot of outliers above the median amount value.

Now we can explore more within parties.
```{r}
#Create candidate datafram
candidate_group <- group_by(FL, party, cand_nm)
FL.contr_by_candidate <- summarize(candidate_group,
                                   sum_candidate = sum(contb_receipt_amt),
                                   mean_can = mean(contb_receipt_amt),
                                   n = n())

FL.contr_by_candidate[rev(order(FL.contr_by_candidate$sum_candidate)),]
```
```{r}
#plot contribution by candidate
ggplot(aes(x = reorder(cand_nm, -sum_candidate/1000), y = sum_candidate/1000), data = FL.contr_by_candidate) +
  geom_bar(aes(fill = party), stat = 'identity') +
  scale_y_continuous(limits = c(0, 23000)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab('Candidate') +
  ylab('Contribution Received (Thousands)') +
  ggtitle('Contribution Received by Candidate') +
  scale_fill_manual(values = c("blue", "red"))
```

```{r}
#Create candidate_party dataframe for percentage calculation
can_party <- left_join(FL.contr_by_candidate, FL.contr_by_party, by = 'party')

can_party
```
```{r}
ggplot(aes(x = reorder(cand_nm, -(sum_candidate/sum_party*100)), y = sum_candidate/sum_party*100), data = can_party) +
  geom_bar(aes(fill = party), stat = 'identity') +
  geom_text(stat='identity', aes(label = paste(round(100*sum_candidate/sum_party,0),'%')), 
            size=3, data = can_party, vjust = -0.4)+
  scale_y_continuous(limits = c(0, 100)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab('Candidate') +
  ylab('Percentage of Donation') +
  ggtitle('Percentage of Contribution Received by Candidate from their Own Party') +
  scale_fill_manual(values = c("blue", 'red'))
```

It's obvious to see that the donation within Democrat party was mainly received by Clinton Hillary, around 86%. For Republican party, Donald Trump received the most contributions with about 38%, Marco Rubio and Jeb Bush received similar amount of contibutions for about 20% and 19%. So for each party, we can also say that the majority of the donations were received by few candidates.

Now let's see is there any pattern of contribution amount by gender, compared with earlier contribution number by gender, which has no obvious pattern.
```{r}
#plot of contribution amount by gender
ggplot(aes(x = gender, y = contb_receipt_amt, fill = gender), 
       data = FL, vjust = -0.4) +
  geom_bar(stat = 'identity') +
  xlab('Gender') +
  ylab('Contribution Amount') +
  ggtitle('Contribution Amout of Contributions by Gender')
```

```{r}
#boxplot of contribution amount by gender
ggplot(aes(x = gender, y = contb_receipt_amt, fill = gender), data = FL) +
  geom_boxplot() +
  xlab('gender') +
  ylab('Contribution Amount') +
  ggtitle('Contribution Amount by Gender Boxplot') +
  coord_cartesian(ylim = c(0, 100))

```
```{r}
by(FL$contb_receipt_amt, FL$gender, summary)
```

Now we see from the contribution amount and median and mean by gender, both plots show male contributions are higher than women contributions. On average male donated 1.5 times more than female.

```{r}
# Top candidates
top_candidate <- c("Clinton, Hillary Rodham", "Trump, Donald J.", "Rubio, Marco", "Bush, Jeb", "Sanders, Bernard")

top_candidate
```

```{r}
# Create gender_to_top_candidate dataframe for bar plot
FL.gen_to_top_candidate <- FL %>%
  filter(FL$cand_nm %in% top_candidate) %>%
  group_by(cand_nm, gender) %>%
  summarize(sum_gen_can = sum(contb_receipt_amt))

FL.gen_to_top_candidate
```

```{r}
# plot of contribution of candidate by gender
ggplot(aes(x = reorder(cand_nm, -(sum_gen_can/1000)), y = sum_gen_can/1000, fill = gender), 
       data = FL.gen_to_top_candidate) +
  geom_bar(stat = 'identity', position = position_dodge(width = 1)) +
  xlab('Candidate') +
  ylab('Contribution Amount (Thousands)') +
  ggtitle('Contribution Amount to Top Candidate by Gender') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Compared with female's support to Donald Trump, Cliton Hillary has higher proportion of female contributions.

Earlier we noticed that the retired people made the most number of contributions, let's see if there is the same pattern when looking into total contribution amount.

```{r}
FL.contr_by_occupation
```
```{r}
ggplot(aes(x = reorder(contbr_occupation, -(sum_occupation/1000)), y = sum_occupation/1000), data = FL.contr_by_occupation) +
  geom_bar(stat = 'identity') +
  geom_text(stat = 'identity', aes(label = round(sum_occupation/1000)), data = FL.contr_by_occupation, vjust = -0.4) +
  xlab('Top 10 Occupations') +
  ylab('Total Contribution Amount (Thousands)') +
  ggtitle('Total Contribution Amount From Top 10 Occupations') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(aes(x = reorder(contbr_occupation, -(mean_occupation)), y = round(mean_occupation,2)), data = FL.contr_by_occupation) + 
  geom_bar(stat = 'identity') +
  geom_text(stat = 'identity', aes(label = round(mean_occupation,2)), data = FL.contr_by_occupation, vjust = -0.4) +
  xlab('Top 10 Occupations') +
  ylab('Average Contribution Amount') +
  ggtitle('Average Contributions From the Top 10 Occupations') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Retired people contributed the most, but when lookint into average contributions, homemaker comes to the first and attorney takes the second place. The unemployed people contribute the least on average among these 10 top occupations.

```{r}
FL.contr_by_city
```


```{r}
#plot of contribution by city
ggplot(aes(x = reorder(contbr_city, -sum_city), y = round(sum_city/1000)), data = FL.contr_by_city) +
  geom_bar(stat = 'identity') +
  xlab('Top 10 City') +
  ylab('Contribution Amounts (K)') +
  ggtitle('Top 10 City Amounts of Donors') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#plot of contribution by city
ggplot(aes(x = reorder(contbr_city, -mean_city), y = round(mean_city)), data = FL.contr_by_city) +
  geom_bar(stat = 'identity') +
  xlab('Top 10 City') +
  ylab('Contribution Amounts') +
  ggtitle('Average of Top 10 City Amounts of Donors') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Miami is still the top cicy in conbributions but if talking about the average, Boca Raton has the highest contibution amount on average.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
* Republican party received almost 1.3 times contribution of Democrat party. But since there were 17 Republican candidates and 5 Democrat candidates, on average Democrat is almost 2.6 times of Republican.
* Hillary Cliton received the most contribution amount, which is 1.73 times of the second position Donald Trump received.
* For each party, the majority of contributions are received by a few candidates.
* Although there are comparable number of contributions for female and male, male donated more than female both in total and on average.
* Retired people contribute the most in total amount, while on average homemaker contribute most, followed by attorney.
* Miami contributes the most in total amount in Florida, but on average Boca Raton contributes the most. 

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
One interesting finding is that on average of the amount, homemaker contributed the most. Although in total their contributions are not the most, and this is because the number of contributions is not much.

### What was the strongest relationship you found?
Since now only contribution amount is numeric variable, we cannot find the correlations.

# Multivariate Plots Section
Now let's explore the top candidates' contribution trend over time. 

```{r}
#Cumulative dataframe
candidate_cum <- FL %>% 
  filter(cand_nm %in% top_candidate) %>% 
  group_by(cand_nm, date) %>% 
  summarize(n = n(), 
            total = sum(contb_receipt_amt)) %>% 
  mutate(cumn = cumsum(n), 
         cum_total = cumsum(total))
```

```{r}
#Cumulative plot (Amount)
ggplot(aes(x = date, y = cum_total/1000, color = cand_nm),
       data = candidate_cum) +
  geom_point(alpha = 1/40) +
  geom_smooth(method = "loess") +
  theme_tufte() +
  xlab("Date") +
  ylab("Cumulative Amount (Thousands)") +
  ggtitle("Cumulative Amount of Top Candidates") +
  theme(plot.title = element_text(size = 16),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "bottom") +
  scale_color_brewer(type = "qual", palette = "Set1")
```

```{r}
#Cumulative plot (number)
ggplot(aes(x = date, y = cumn/1000, color = cand_nm),
       data = candidate_cum) +
  geom_point(alpha = 1/40) +
  geom_smooth(method = "loess") +
  theme_tufte() +
  xlab("Date") +
  ylab("Cumulative Count") +
  ggtitle("Cumulative Count of Top Candidates") +
  theme(plot.title = element_text(size = 16),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "bottom") +
  scale_color_brewer(type = "qual", palette = "Set1")
```

We can notice that Cliton Hillary and Donald Trump both have steep trend in cumulative donation amount and count, but the thrend of Donald Trump are both under Cliton Hillary's trends, that means Cliton Hillary received higher contributions earlier than Donald Trump. Bernard Sanders's donation count trend is even more steep than Cliton Hillary but the donation amount trend is not, this tell us most supporters for Bernard Sanders are smaller donors.

```{r }
# Create a new dataframe for map
lat <- FL$latitude
lon <- FL$longitude
party <- FL$party
FL_map <- data.frame(party, lat, lon)
colnames(FL_map) <- c('party', 'lat', 'lon')

# make_bbox function to get a zoom level
sbbox <- make_bbox(lon = FL$lon, lat = FL$lat, f = 0.01)

my_map <- get_map(location = sbbox, maptype = "roadmap", scale = 2, color="bw", zoom = 7)

ggmap(my_map) +
  geom_point(data=FL_map, aes(x = lon, y = lat, color = party), 
             size = 2, alpha = 0.2) +
  xlab('Longitude') +
  ylab('Latitude') +
  ggtitle('Location of Donors') +
  scale_y_continuous(limits = c(41, 43)) +
  scale_x_continuous(limits = c(-74, -70)) +
  scale_color_manual(breaks=c("democrat", "republican"), values=c("blue","red"))
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!