Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('C:/Files/Udacity/Data-Analysis/Notes/Exploratory Data Analysis (R)/R Basics/eda-course-materials/lesson3/pseudo_facebook.tsv', sep = '\t')
#sep="\t" tells R that the file is tab-delimited (use " " for space delimited and "," for comma delimited; use "," for a .csv file)

qplot(x = age, y = friend_count, data = pf)
#qplot(age, friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf)  + 
  geom_point(alpha = 1/20) + 
  xlim(13, 90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf)  + 
  geom_jitter(alpha = 1/20) + 
  xlim(13, 90)

summary(pf$age)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!
```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf)  + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```


#### What do you notice?

***

### Alpha and Jitter
Notes:
```{r Alpha and Jitter}
#Explore the relationship between friends initiated(y) vs age(x).
ggplot(aes(x = age, y = friendships_initiated), data = pf)  + 
  geom_jitter(alpha = 1/10, position = position_jitter(h = 0)) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)
```

```{r}
age_groups <- group_by(pf, age)
pf.fc_by_age<-summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>%
  arrange(age)

head(pf.fc_by_age)
```

```{r}
# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.

ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age)  + 
  geom_line()
```


***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  #xlim(13, 90) +    replace by coord_cartesian
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) +
  geom_point(alpha = 0.05,
             position = position_jitter(h = 0),
             color = 'orange') +
  #coord_trans(y = 'sqrt') +  replace by coord_cartesian
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .1), 
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .5), 
            color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .9), 
            linetype = 2, color = 'blue')
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = 'pearson')
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r}
with(pf, cor.test(age, friend_count, method = 'pearson'))
```

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count), method = 'pearson')
#method = 'pearson' can be removed since cor.test default method is 'pearson'
```



***

### Correlation Methods
Notes:
```{r}
with(subset(pf, age <= 70), cor.test(age, friend_count), method = 'spearman')
```

***

## Create Scatterplots
Notes:

```{r}
# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x). Use any of the
# techniques that you've learned so far to
# modify the plot.
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point()
```

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')
```


***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$www_likes_received, pf$likes_received)
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

```{r}
data(Mitchell)
```


Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()

qplot(data = Mitchell, Month, Temp)
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(data = Mitchell, aes(x = Month, y = Temp)) +
  geom_point() +
  scale_x_discrete(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

```{r}
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()
```
```{r}
install.packages("devtools")
library(devtools)
```


```{r}
x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
#dcor.ttest(x, y)
```


***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()

head(pf.fc_by_age)
```

```{r}
# Create a new variable, 'age_with_months', in the 'pf' data frame.
# Be sure to save the variable in the data frame rather than creating
# a separate, stand-alone variable. You will need to use the variables
# 'age' and 'dob_month' to create the variable 'age_with_months'.

# Assume the reference date for calculating age is December 31, 2013.
pf$age_with_months <- pf$age + (12 - pf$dob_month)/12 
```


***

### Age with Months Means

```{r Age with Months Means}
# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data framed should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean	friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11

```

Programming Assignment
```{r Programming Assignment}
library(dplyr)
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
           friend_count_median = median(friend_count),
           n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

```{r Alternate Solution}
age_with_months_groups <- gropu_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_with_months)
```

```{r}
age_with_months_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months2 <- summarise(age_with_months_groups,
                                  friend_count_mean = mean(friend_count),
                                  friend_count_median = median(friend_count),
                                  n = n())

pf.fc_by_age_months2 <- arrange(pf.fc_by_age_months2, age_with_months)

head(pf.fc_by_age_months2)
```


***

### Noise in Conditional Means

```{r Noise in Conditional Means}
# This programming assignment will not be graded, but when you submit your code,
# the assignment will be marked as correct. By submitting your code, we can add
# to the feedback messages and address common mistakes in the Instructor Notes.

ggplot(aes(x = age_with_months, y = friend_count_mean),
       data = subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean),
             data = subset(pf.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean),
             data = subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count),
             data = subset(pf, age < 71)) +
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

