---
title: "Lesson 4 Quiz"
output: html_document
---
##Load Diamond Data##
```{r}
library(ggplot2)     #must load the ggplot package first
data(diamonds)       #loads the diamonds data set since it comes with the ggplot package
summary(diamonds)
```

##Price Histogram##
# Create a histogram of the price of
# all the diamonds in the diamond data set.
```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(binwidth = 100) +
  scale_x_continuous(breaks = seq(0, 18823, 1000))  #long_tailed should be added
```


##Price Histogram Summary##
```{r}
subset(diamonds, price < 500)
```

```{r}
subset(diamonds, price < 250)
```

```{r}
subset(diamonds, price >= 15000, select=c(price))
```

##Cheaper Diamonds##
# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won't be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Submit your final code when you are ready.
```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(binwidth = 25, color = 'black', fill = '#099DD9') +
  scale_x_continuous(breaks = seq(0, 1500, 100), limits = c(0,1500))
ggsave('priceHistogram.png')
```

##Price by Cut Histograms##
# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.
```{r}
ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 100, 
                                  color = "black", fill = "#099DD9") + 
  ggtitle("Diamonds Price by Cut Histogram") + 
  facet_grid(. ~ cut)
```

##Price by Cut##
#Which cut has the highest priced diamond?
#Which cut has the lowest priced diamond?
#Which cut has the lowest median price?
```{r}
by(diamonds$price,diamonds$cut,summary)
```


