Lesson 6
========================================================

### Welcome
Notes:

***

### Scatterplot Review

```{r Scatterplot Review}
# Let's start by examining two variables in the data set.
# The scatterplot is a powerful tool to help you understand
# the relationship between two continuous variables.

# We can quickly see if the relationship is linear or not.
# In this case, we can use a variety of diamond
# characteristics to help us figure out whether
# the price advertised for any given diamond is 
# reasonable or a rip-off.

# Let's consider the price of a diamond and it's carat weight.
# Create a scatterplot of price (y) vs carat weight (x).

# Limit the x-axis and y-axis to omit the top 1% of values.

qplot(data = diamonds, x = carat, y = price,
      xlim = c(0, quantile(diamonds$carat, 0.99)),
      ylim = c(0, quantile(diamonds$price, 0.99))) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21)
```

```{r}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point(color = '#F79420', alpha = 1/4) +
  stat_smooth(method = 'lm') +
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, 0.99))) +
  scale_y_continuous(lim = c(0, quantile(diamonds$price, 0.99)))
```

***

### Price and Carat Relationship
Response:
A nonlinear relationship. Maybe it's exponential. Also the dispersion or variance or the relationship also increase as carat size increases. 
***

### Frances Gerety
Notes:

#### A diamonds is


***

### The Rise of Diamonds
Notes:

***

### ggpairs Function
Notes:
```{r}
# install these if necessary
install.packages('GGally')  #particular plot 
install.packages('scales')  #for a variety of things
install.packages('memisc')  #summarize the regression
install.packages('lattice') #few other things
install.packages('MASS')    #various functions
install.packages('car')     #recode variables
install.packages('reshape') #reshape and wrangle data
install.packages('dplyr')   #create summaries and transmissions
```

```{r}
# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
```

```{r ggpairs Function}
# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, params = c(shape = I('.'), outlier.shape = I('.')))
```

What are some things you notice in the ggpairs output?
Response:
There may be relationships between price and clarity and price and color.
And the critical factor dribing price is the size, or the carat weight of the diamond, it is nonlinear, even exponential.
***

### The Demand of Diamonds
Notes:

```{r The Demand of Diamonds}
# Create two histograms of the price variable
# and place them side by side on one output image.

# We've put some code below to get you started.

# The first plot should be a histogram of price
# and the second plot should transform
# the price variable using log10.

# Set appropriate bin widths for each plot.
# ggtitle() will add a title to each histogram.

library(gridExtra)

plot1 <- qplot(data = diamonds, x = price, binwidth = 100, fill = I('#0990D9')) + 
  ggtitle('Price')

plot2 <- qplot(data = diamonds, x = price, binwidth = 0.01, fill = I('#F79420')) +
  ggtitle('Price (log10)') +
  scale_x_log10()

library(gridExtra)
library(grid)
grid.arrange(plot1, plot2, ncol = 2)
```

***

### Connecting Demand and Price Distributions
Notes:
The price of diamonds are pretty heavily skewed, but when add price log function, it is much like a normal distribution, and can even see a little bit of bimodality on the log plot. 
***

### Scatterplot Transformation

```{r Scatterplot Transformation}
qplot(carat, price, data = diamonds) +
  scale_y_continuous(trans = log10_trans()) +
  ggtitle('Price (log10) by Carat')
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat), decreasing = T))
head(sort(table(diamonds$price), decreasing = T))
```


```{r Overplotting Revisited}
# Add a layer to adjust the features of the
# scatterplot. Set the transparency to one half,
# the size to three-fourths, and jitter the points.

ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 0.75, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:

***

### Price vs. Carat and Clarity

Alter the code below.

```{r}
# install and load the RColorBrewer package
install.packages('RColorBrewer', dependencies = TRUE)
library(RColorBrewer)
```

```{r Price vs. Carat and Clarity}
# Adjust the code below to color the points by clarity.

# A layer called scale_color_brewer() has 
# been added to adjust the legend and
# provide custom colors.
library(scales)

ggplot(aes(x = carat, y = price, colour = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:
Diamonds with lower clarity are almost always cheaper than that with better clarity. 
***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
library(scales)

ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response:
Most of the data with ideal cut, so we lost the cut pattern.
***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = FALSE,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response:

***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
Notes:

```{r Building the Linear Model}
library(memisc)

m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
#I stands for as is
#This is instead of instrucing R to interpret these symbols as part of the formula to construct the design matrix for the regression.
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5, sdigits = 3)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)
Our data starts from 2008, so we need to account for inflation, but the diamond market is quite different now than it was. And in fact when fitting models to this data and predicted the price of the diamonds that found off a market, it kept getting predictions that were way too low. After some additional digging, and that is because that global diamonds were poor. It turns out that prices plummeted in 2008 due to the global financial crisis. And since then prices, at least for wholesale polished diamonds, have grown at about 6% per year, compound annual rate. The rapidly growing number of couples in China buying diamond engagement rings might also explain this increase. And the diamond prices grew unevenly across different karat sizes since 2008, means the modle estimated couldn't simply be adjusted by inflation. 
Response:

***

### A Bigger, Better Data Set
Notes:
```{r}
#install bitops and RCurl
install.packages('bitops')
install.packages('RCurl')
library('bitops')
library('RCurl')
```


```{r A Bigger, Better Data Set}
#diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
#load(rawConnection(diamondsurl))

load("C:/Files/Udacity/Data-Analysis/Notes/Exploratory Data Analysis (R)/R Basics/eda-course-materials/lesson6/BigDiamonds.rda")
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}
#We will only use GIA certified diamonds in this modle, and only look at diamonds under $10,000 because these are the type of diamonds sold at most retailers and hence the kind we care most about.
#By trimming the most expensive diamonds from the data set, our model will also be less likely to be thrown off by outliers and the higher variants at the high-end of price and carat.

# Your task is to build five linear models like Solomon
# did for the diamonds data set only this
# time you'll use a sample of diamonds from the
# diamondsbig data set.

# Be sure to make use of the same variables
# (logprice, carat, etc.) and model
# names (m1, m2, m3, m4, m5).

# To get the diamondsbig data into RStudio
# on your machine, copy, paste, and run the
# code in the Instructor Notes. There's
# 598,024 diamonds in this data set!

# Since the data set is so large,
# you are going to use a sample of the
# data set to compute the models. You can use
# the entire data set on your machine which
# will produce slightly different coefficients
# and statistics for the models.

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamondbig_samp <- diamondsbig[sample(1:length(diamonds$price), 10000), ]
```

```{r}
summary(diamondbig_samp)
```

```{r}
#diamondbig_samp$logprice = log(diamondbig_samp$price)
m1 <- lm(I(log(price)) ~ I(carat^(1/3)),
         data = diamondbig_samp[diamondbig_samp$price < 10000 &
                                  diamondbig_samp$cert == "GIA",])
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5, sdigits = 3)
```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601
```{r}
library(memisc)
```

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

```{r}
install.packages("zipcode")
```
```{r}
install.packages("gender")
```

```{r}
install.packages("choroplethr")
```
```{r}
install.packages("choroplethrMaps")
```

```{r}
install.packages("chroplethrZip")
```
```{r}
install.packages("choroplethrZip")
```
```{r}
install.packages("devtools")
```
```{r}
library(devtools)
install_github('arilamstein/choroplethrZip@v1.4.0')
```

